---
title: "Impact of Euthanasia on Suicide Rates among Elderly"
subtitle: "Empirical Evidence from European Regions"
author:
  - Wenjie Tu
  - Christian Birchler
date: "Spring Semester 2021"
output: pdf_document
---

```{r echo=F}
Sys.setenv(LANG="en")
rm(list=ls())
setwd('F:/University of Zurich/Second Semester/Program Evaluation and Causal Inference/suicide')
```

```{r message=F, warning=F}
# import libraries
library(ggplot2)
library(stargazer)
library(tidyverse)
library(plm)
library(Synth)
library(gsynth)
library(panelView)
```

```{r}
# read data
dta <- read.csv('./References/hlth_cd_acdr_1_Data.csv')
```

### Data Preparation

```{r}
# data types of columns
str(dta)
```

```{r}
# drop columns
dta <- dta[, !names(dta) %in% c('UNIT', 'SEX', 'AGE')]

# rename columns
colnames(dta) <- c('year', 'region', 'dth.cause', 'dth.rate')

# create a variable for country code
dta$country <- substr(dta$region, 1, 2)
```

```{r}
# drop countries
dta <- dta[!dta$country %in% c('HU', 'AT', 'CZ', 'DE'), ]
```

Reasons for dropping these countries:

* We included these countries separately or all at once and found that the estimated missing counterfactuals do not fit the observed outcomes very well in pre-treatment periods in the plot.
* We also looked at regional suicide rates in these countries separately and found that regional observations vary a lot even within the same country and the aggregate suicide rates (at country levels) are higher than other countries. We assume that there exist some unobservable region-specific interventions.
* We aggregated regional suicide rates to country levels and there is no (quasi-) parallel trend in country-level suicide rates between these countries and Netherlands.

We therefore conclude that including these countries all at once or separately would bias the missing counterfactual estimation in some sense and it is better to drop them.

```{r}
# subset data by cause of death
d.suicide <- dta[dta$dth.cause=='Intentional self-harm', ]

d.suicide <- subset(d.suicide, select=-c(dth.cause))
```

```{r}
# create a dummy variable to indicate the time when the treatment started
d.suicide$time <- ifelse(d.suicide$year >= 2002, 1, 0)

# create a dummy variable to identify the group exposed to the treatment
d.suicide$treated <- ifelse(d.suicide$country == 'NL', 1, 0)

# create an interaction between time and treated.
d.suicide$did <- d.suicide$time*d.suicide$treated
```

```{r}
# create a numerical variable for region codes
d.suicide$unit.num <- factor(
  d.suicide$region, 
  levels = unique(d.suicide$region), 
  labels = seq(length(unique(d.suicide$region)))
)

d.suicide$unit.num <- as.numeric(d.suicide$unit.num)

d.suicide <- as.data.frame(d.suicide)

str(d.suicide)
```

### Data Visualization

```{r}
# visualize the data structure
panelView(dth.rate ~ did, data=d.suicide, cex.axis=2, 
          index = c('unit.num', 'year'), pre.post=T)
```

```{r}
# visualize the outcome variable
# different colors correspond to different treatment status.
panelView(dth.rate~did, data=d.suicide, index=c('unit.num', 'year'), 
          type='outcome', ylab='crude suicide rates', 
          main='Legalization of Euthanasia and Crude Suicide Rates')
```

### Data Analysis

```{r message=F}
set.seed(2021)
out <- gsynth(dth.rate ~ did, data = d.suicide, 
              EM = TRUE, # Expectation Maximization algorithm will be used
              index = c("unit.num","year"), # specify group and time variables
              force='two-way', # two-way fixed effects
              CV = TRUE, # cross validation
              r = c(0, 5), # the number of factors
              se = TRUE, # uncertainty estimates will be produced
              inference = 'parametric', # inferential method
              nboots = 1000, # the number of bootstrap runs
              parallel = TRUE, # parallel computing
              cores = 5, # the number of cores used in parallel computing
              seed = 2021 # set seed for reproducibility
              )
```

* **sigma2** stands for the estimated variance of the error term.
* **IC** represents the Bayesian Information Criterion.
* **MPSE** is the Mean Squared Prediction Error.

The cross-validation procedure selects an $r$ that minimizes the MSPE. When $r=0$ (no factor is assumed), the estimates are close to what we obtain from difference-in-differences.

```{r message=F, warning=F}
plot(out, main='Estimated ATT', ylab="estimated ATT")

plot(out, type = "counterfactual", ylab="crude suicide rate", 
     raw = "none", main="Treated and Counterfactual Averages")
```

In the plot of **Estimated ATT**, the estimated ATT is centered around zero horizontal line in pre-treatment periods and is statistically different from zero in post-treatment periods.

In the plot of **Treated and Counterfactual Averages**, the estimated curve for missing counterfactuals fits the curve for observed outcome averages quite well in pre-treatment periods and they demonstrate different patterns in post-treatment periods. The plot is consistent with the criteria for synthetic control group.

```{r message=F, results='asis'}
stargazer(out$est.att, header=F, title="ATT by period", type = "text")

stargazer(out$est.avg, header=F, type = "text", title="ATT averaged over all periods")
```

```{r}
d.suicide <- pdata.frame(d.suicide, index = c("unit.num", "year"))
```

```{r warning=F, message=F}
didreg.ols <- lm(dth.rate ~ treated + time + did, data = d.suicide)

didreg.fe <- plm(dth.rate ~ treated + time + did, data = d.suicide, 
                 effect = "twoways", model = "within")
stargazer(didreg.ols, didreg.fe, header = F, type = "text", 
          keep = "did", dep.var.caption = "suicide rate", 
          covariate.labels = "ATT.avg", single.row = T, keep.stat = "n", no.space = T, 
          column.labels = c("pooling", "two-way FE"), dep.var.labels.include = F, 
          title = "ATT averaged without using GSC")
```